---
title: "LAGOS Spatial Analysis"
author: "Matthew Ross"
date: "9/11/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
library(ggplot2)  #for graphing data
```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)
```{r data-read}
# #Lagos download script
LAGOSNE::lagosne_get(dest_folder = LAGOSNE:::lagos_path())


#Load in lagos
lagos <- lagosne_load()

#Grab the lake centroid info
lake_centers <- lagos$locus

# This chunk returns error, 'no such file or directory', does not seem important
# load('lake_centers.Rdata')

```



### Convert to spatial data
```{r}
#Look at the column names
#names(lake_centers)

#Look at the structure
#str(lake_centers)

#View the full dataset
# View(lake_centers %>% slice(1:100))

spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
                st_transform(2163)

#Subset for plotting
subset_spatial <- spatial_lakes %>%
  slice(1:100)

#Dynamic mapviewer
mapview(subset_spatial)

```


### Subset to only Minnesota

```{r}
states <- us_states()

#Plot all the states to check if they loaded
#mapview(states)
minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(2163)

#Subset lakes based on spatial position
minnesota_lakes <- spatial_lakes[minnesota,]

#Plotting the first 1000 lakes
minnesota_lakes %>%
  slice(1:1000) %>%
  arrange(lake_area_ha) %>%
  mapview(.,zcol = 'lake_area_ha')
```



# In-Class work


## 1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)

```{r}
ill_iowa <- states %>%
  filter(name %in% c('Iowa','Illinois')) %>%
  st_transform(2163)

mapview(ill_iowa)

```


## 2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa combined? How does this compare to Minnesota?

```{r}
#Subset lakes only in iowa and illinois
ill_iowa_lakes <- spatial_lakes[ill_iowa,]

#combined Illinois and Iowa have 16466 lakes while minnesota has 29038. This is determined by looking at the number of observations in the dataframes.

```


## 3) What is the distribution of lake size in Iowa vs. Minnesota?

- Here I want to see a histogram plot with lake size on x-axis and frequency on y axis (check out geom_histogram)

```{r}
#Subset lakes only in iowa. 
# Separate Iowa from other states
iowa <- states %>%
  filter(name %in% c('Iowa')) %>%
  st_transform(2163)
#Select only lakes in iowa
iowa_lakes <- spatial_lakes[iowa,]%>%
  arrange(lake_area_ha)

# Used to understand the size distribution to create the histograms
# max(iowa_lakes$lake_area_ha)
# min(iowa_lakes$lake_area_ha)

#Iowa Histogram Lake size vs frequency
ggplot(data=iowa_lakes, aes(iowa_lakes$lake_area_ha)) +     
    geom_histogram(breaks = seq(0,5000, by = 50))+
    scale_x_log10()+
    labs(title="Histogram for Iowa") +
    labs(x="Lake Area", y="Count")

# Used to understand the size distribution to create the histograms
# arrmin_lake<-minnesota_lakes %>%
#   arrange(lake_area_ha)

# max(minnesota_lakes$lake_area_ha)
# min(minnesota_lakes$lake_area_ha)

#Minnesota Histogram Lake size vs frequency
ggplot(data=minnesota_lakes, aes(minnesota_lakes$lake_area_ha))+      
    geom_histogram(breaks = seq(0,124000, by = 1000))+
    scale_x_log10()+
    labs(title="Histogram for Minnesota") +
    labs(x="Lake Area", y="Count")

```


## 4) Make an interactive plot of lakes in Iowa and Illinois and color them
by lake area in hectares

```{r}
ill_iowa_lakes %>%
  arrange(lake_area_ha) %>%
  mapview(.,zcol = 'lake_area_ha')
```


## 5) What other data sources might we use to understand how reservoirs and 
natural lakes vary in size in these three states? 
# It would be interesting to look at lake perimeter and possibly make a relationship between perimeter and area. This could give us more information on what kind of habitat these lakes are providing. Additionally, we could look at the area of the state and calculate the percentage that is covered in lakes.

